Type-Safe Modular Parsing

### 2017-09-05 SLE_17/sections/ObjectAlgebras.tex
Under 12 pages.

\vspace{-4pt}
\lstinputlisting[linerange=5-8]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_ALG
\vspace{-4pt}
Here \inlinecode{Alg} is called an \textit{Object Algebra interface}, parameterized by the type
\inlinecode{E}, which abstracts over the concrete type of the AST.

@@

\vspace{-4pt}
\lstinputlisting[linerange=12-16]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_PRINT
\vspace{-4pt}
Here \inlinecode{Print} is called an \textit{Object Algebra}. It traverses an expression bottom-up, and returns a string as the result.
One can also define an evaluation operation as a new trait that extends \inlinecode{Alg[Int]}. Hence adding new operations is modular. We omit that code
due to space reasons.

%\lstinputlisting[linerange=20-23]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_EVAL

@@

\vspace{-4pt}
\lstinputlisting[linerange=27-29]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_ALGEXT
\vspace{-4pt}
Now pretty-printing on the new language can be realized without modifying existing code:

\vspace{-4pt}
\lstinputlisting[linerange=33-35]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_EXTPRINT
\vspace{-4pt}
An observation is that only the new case is implemented for pretty-printing, and the others have been inherited.
Thus existing code was reused and was not modified!

@@

\vspace{-4pt}
\lstinputlisting[linerange=42-43]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_MAKEEXP
\vspace{-4pt}

@@

\vspace{-4pt}
\lstinputlisting[linerange=8-17]{code/src/papercode/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER_BAD
\vspace{-4pt}
Such a parser looks fine, but it is not extensible. For example, we have demonstrated in Section~\ref{sec:inheritance} that method overriding is essential to update \inlinecode{pExpr} for an extended syntax. However, trying to do a similar method overriding for \inlinecode{pExpr} would require a type \inlinecode{VarAlg[E] =>} \inlinecode{Parser[E]}, which is a \emph{supertype} of the old type \inlinecode{Alg[E] =>} \inlinecode{Parser[E]}, since
the \emph{extended} Object Algebra interface appears in \emph{contravariant} position. This violates
overriding in Scala.

@@

\begin{figure}
  \lstinputlisting[linerange=21-30]{code/src/papercode/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER
  \caption{Pattern of modular parsing using Object Algebras.}
  \label{fig:oa-parser}
  \vspace{-0.2cm}
\end{figure}

@@

\vspace{-4pt}
\lstinputlisting[linerange=34-39]{code/src/papercode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER
\vspace{-4pt}

@@

\vspace{-4pt}
\lstinputlisting[linerange=46-49]{code/src/papercode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER_CLIENT
\vspace{-4pt}

>>>>

\vspace{-4pt}
\lstinputlisting[linerange=5-7]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_ALG
\vspace{-4pt}
Here \inlinecode{Alg} is called an \textit{Object Algebra interface}, parameterized by the type
\inlinecode{E}, which abstracts over the concrete type of the AST.

@@

\vspace{-4pt}
\lstinputlisting[linerange=11-14]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_PRINT
\vspace{-4pt}
Here \inlinecode{Print} is called an \textit{Object Algebra}. It traverses an expression bottom-up, and returns a string as the result.
One can also define an evaluation operation as a new trait that extends \inlinecode{Alg[Int]}. Hence adding new operations is modular. We omit that code
due to space reasons.

%\lstinputlisting[linerange=18-20]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_EVAL

@@

\vspace{-4pt}
\lstinputlisting[linerange=24-25]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_ALGEXT
\vspace{-4pt}
Now pretty-printing on the new language can be realized without modifying existing code:

\vspace{-4pt}
\lstinputlisting[linerange=29-30]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_EXTPRINT
\vspace{-4pt}
An observation is that only the new case is implemented for pretty-printing, and the others have been inherited.
Thus existing code was reused and was not modified!

@@

\vspace{-4pt}
\lstinputlisting[linerange=37-38]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_MAKEEXP
\vspace{-4pt}

@@

\vspace{-4pt}
\lstinputlisting[linerange=8-16]{code/src/papercode/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER_BAD
\vspace{-4pt}
Such a parser looks fine, but it is not extensible. For example, we have demonstrated in Section~\ref{sec:inheritance} that method overriding is essential to update \inlinecode{pExpr} for an extended syntax. However, trying to do a similar method overriding for \inlinecode{pExpr} would require a type \inlinecode{VarAlg[E] =>} \inlinecode{Parser[E]}, which is a \emph{supertype} of the old type \inlinecode{Alg[E] =>} \inlinecode{Parser[E]}, since
the \emph{extended} Object Algebra interface appears in \emph{contravariant} position. This violates
overriding in Scala.

@@

\begin{figure}
  \lstinputlisting[linerange=20-28]{code/src/papercode/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER
  \caption{Pattern of modular parsing using Object Algebras.}
  \label{fig:oa-parser}
  \vspace{-0.2cm}
\end{figure}

@@

\vspace{-4pt}
\lstinputlisting[linerange=32-36]{code/src/papercode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER
\vspace{-4pt}

@@

\vspace{-4pt}
\lstinputlisting[linerange=43-46]{code/src/papercode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER_CLIENT
\vspace{-4pt}

### 2017-09-05 SLE_17/sections/PackratParsing.tex
Shorten more.
Under 12 pages.

\begin{itemize}[leftmargin=*]

@@

\end{itemize}

@@

\vspace{-4pt}
\lstinputlisting[linerange=9-15]{code/src/papercode/Sec2Packrat/Code1.scala}% APPLY:linerange=PACKRAT_OO_AST
\vspace{-4pt}

@@

\vspace{-4pt}
\lstinputlisting[linerange=19-26]{code/src/papercode/Sec2Packrat/Code1.scala}% APPLY:linerange=PACKRAT_SIMPLE_EXPR
\vspace{-4pt}

@@

\begin{table}
\begin{tabularx}{\linewidth}{X}
\hline
\lstinline{def ~[U](q: => Parser[U]): Parser[~[T, U]]} \\
- A parser combinator for sequential composition. \\ \hline
\lstinline{def \^\^[U](f: (T) => U): Parser[U]} \\
- A parser combinator for function application. \\ \hline
%\lstinline{def \^\^\^[U](v: => U): Parser[U]} \\
%- A parser combinator that changes a successful result into the specified value. \\ \hline
\lstinline{def <~[U](q: => Parser[U]): Parser[T]} \\
- A parser combinator for sequential composition which keeps only the left result. \\ \hline
\lstinline{def ~>[U](q: => Parser[U]): Parser[U]} \\
- A parser combinator for sequential composition which keeps only the right result. \\ \hline
\lstinline{def ident: Parser[String]} \\
- A parser which matches an identifier. \\ \hline
\lstinline{def numericLit: Parser[String]} \\
- A parser which matches a numeric literal. \\ \hline
\lstinline{def |[U >: T](q: => Parser[U]): Parser[U]} \\
- A parser combinator for alternative composition. \\ \hline
\lstinline{def |||[U >: T](q0: => Parser[U]): Parser[U]} \\
- A parser combinator for alternative with longest match composition. \\ \hline
\end{tabularx}
\caption{Common combinators from the Scala standard parser combinator library.}\label{tab:packrat}
\end{table}

>>>>

\vspace{-2pt}
\begin{itemize}[leftmargin=*]

@@

\end{itemize}
\vspace{-2pt}

@@

\vspace{-4pt}
\lstinputlisting[linerange=9-13]{code/src/papercode/Sec2Packrat/Code1.scala}% APPLY:linerange=PACKRAT_OO_AST
\vspace{-4pt}

@@

\vspace{-4pt}
\lstinputlisting[linerange=17-24]{code/src/papercode/Sec2Packrat/Code1.scala}% APPLY:linerange=PACKRAT_SIMPLE_EXPR
\vspace{-4pt}

@@

\begin{table}
\begin{tabularx}{\linewidth}{X}
\hline
\lstinline{def ~[U](q: => Parser[U]): Parser[~[T, U]]} \\
- A parser combinator for sequential composition. \\ \hline
\lstinline{def \^\^[U](f: (T) => U): Parser[U]} \\
- A parser combinator for function application. \\ \hline
%\lstinline{def \^\^\^[U](v: => U): Parser[U]} \\
%- A parser combinator that changes a successful result into the specified value. \\ \hline
\lstinline{def <~[U](q: => Parser[U]): Parser[T]} \\
- A parser combinator for sequential composition which keeps only the left result. \\ \hline
\lstinline{def ~>[U](q: => Parser[U]): Parser[U]} \\
- A parser combinator for sequential composition which keeps only the right result. \\ \hline
\lstinline{def ident: Parser[String]} \\
- A parser which matches an identifier. \\ \hline
\lstinline{def |[U >: T](q: => Parser[U]): Parser[U]} \\
- A parser combinator for alternative composition. \\ \hline
\lstinline{def |||[U >: T](q0: => Parser[U]): Parser[U]} \\
- A parser combinator for alternative with longest match composition. \\ \hline
\end{tabularx}
\caption{Common combinators from the Scala standard parser combinator library.}\label{tab:packrat}
\end{table}
